aRianna: R package for consistency scores in text data

### 2020-09-01 outputs/documents/paper_ngrams.Rmd
ngram paper draft ready for revie

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    toc: false
    fig_caption: true
    latex_engine: pdflatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
biblio-style: apalike
title: "Consistency scores in text data"
thanks: "Thank you to **X, Y and Z** for helpful comments. A Shiny application for interactive and small-scale examples is available: https://kelichiu.shinyapps.io/Arianna/. An R package for larger scale applications is available: **HERE**. Our code and datasets are available: https://github.com/RohanAlexander/consistency_scores_in_text_datasets. Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Ke-Li Chiu
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: University of Toronto
abstract: "In this paper we introduce a process to clean the text extracted from PDFs using n-gram models that is widely used in statistical natural language processing. Our approach compares originally extracted text with the text generated from, or expected by, these methods using earlier text as stimulus. To guide this process, we introduce the notion of a consistency score, which refers to the proportion of text that is unchanged by the model. This is used to monitor changes during the cleaning process, and to compare the messiness of different texts. We illustrate our process on text from the book Jane Eyre and introduce both a Shiny application and an R package to make our process easier for others to adopt."
keywords: "text-as-data; natural language processing; quantitative analysis."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
# spacing: double
endnote: no
graphics: yes
---

@@

# Introduction

@@

We employ n-gram models to calculate consistency scores and generate word candidates for text corrections. The n-gram approach involves identifying two, three, or more, words that are commonly found together. **(Keli adds descriptions on how n-gram can be used to calculate consistency scores and generate word candidates)**

The remainder of our paper is structured as follows: **(Keli add structure)**. Additionally, we construct a Shiny app available at: https://kelichiu.shinyapps.io/aRianna/. That app computes internal and external consistency scores for corpus excerpts, and have developed an R Package, `aRianna`, that allows our approach to be used on larger datasets, which is available at: https://github.com/RohanAlexander/aRianna.

# Background

@@

Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class]. In our application, we use a tri-gram based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability.

### Package Dependencies

There are a variety of ways to implement an n-gram model within R @citeR including using R packages such as `Quanteda` [@citequanteda], `tidyText` [@citetidytext] and `tm` [@citetm]. In our package, we used the `Quanteda` R package because it has a comprehensive set of functions for conducting text analysis. We also employed `dplyr` (**Keli adds citation**) for data frame manipulation and `tibble` (**Keli adds citation**) for transforming data frames to the tibble format:

@@

### make_internal_consistency_dataset Function

To obtain the internal consistency score, we will construct a internal consistency dataset from a body of text that is the source of the text being evaluated. The function `make_internal_consistency_dataset` is created for this purpose. The function turns the body_of_text into an internal consistency dataset which the text being evaluated is compared with in the following steps. To do so, the function first transforms the body of text into word tokens with the punctuation removed and the letters turned to lowercase:

@@

Finally, the `make_internal_consistency_dataset` function splits the tri-grams are to two parts: the previous two words and the last word:

@@

### generate_internal_consistency_score Function

After we have the internal consistency dataset, we need a function to compare the text being evaluated and the internal dataset in order to retrieve the internal consistency score — the `generate_internal_consistency_score` function. Similar to the `make_internal_consistency_dataset` function, the function first transforms the text being evaluated into tokens, then split tri-grams and preserve them into a tibble:

@@

Next, the function combines the text tokens tibble and the internal consistency tibble by the left_join function of `dplyr`. The left_join function returns all the rows from the text tokens tibble and all the columns from both tibbles. By combining the two tibbles, the function generates two addtional columns — the `last_word` column are the last words of all tri-grams, and the `last_word_expected` contains the words that are in the internal consistency dataset. By comparing the words in `last_word` and `last_word_expected`. If a last word is in `last_word` but not `last_word_expected`, it is regarded as an unexpected word and the corresponding word in `last_word_expected` serves as a replacement candidate:

@@

### generate_internal_consistency_score Function

The `generate_external_consistency_score` function works identically with the `generate_internal_consistency_score`. The only difference is that it compares the text being evaluated with a external consistency dataset that is larger and more general. The external consistency dataset is from **(Keli adds source of the external dataset)**.

### Demonstration

@@

```{r, eval=FALSE}
aRianna::generate_internal_consistency_score(text_to_evaluate, internal_consistency_dataset)

@@

## $`unexpected words`
## # A tibble: 13 x 4
##    first_words last_word last_word_expected as_expected
##    <chr>       <chr>     <chr>              <lgl>      
##  1 there_was   na        a                  FALSE      
##  2 there_was   na        an                 FALSE      
##  3 there_was   na        no                 FALSE      
##  4 there_was   na        not                FALSE      
##  5 there_was   na        nothing            FALSE      
##  6 there_was   na        of                 FALSE      
##  7 there_was   na        one                FALSE      
##  8 there_was   na        only               FALSE      
##  9 there_was   na        plenty             FALSE      
## 10 there_was   na        someone            FALSE      
## 11 there_was   na        something          FALSE

@@

# Stylized example

As a stylized example, let's consider the following actual paragraph from Jane Eyre, by Charlotte Bronte:

> There was no possibility of taking a walk that day. We had been wandering, indeed, in the leafless shrubbery an hour in the morning; but since dinner (Mrs. Reed, when there was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, that further out-door exercise was now out of the question.

Let's pretend that this text had been created from optical character recognition and that it had the following errors: some 'h' were replaced with 'b'; and some 'd' have been replaced with 'al':

> There was no possibility of taking a walk that day. We had been wandering, indeed, in tbe leafless shrubbery an hour in tbe morning; but since dinner (Mrs. Reeal, when tbere was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, tbat further out-door exercise was now out of the question.

Assume a model that is trained to perfectly forecast the next word in Jane Eyre. For this fragment there are 62 words, comprising 5 errors and 57 correct words. So the internal consistency score of this fragment would be: 57/62 = 0.919. When we recognise and correct the errors, this consistency score would increase to 1.

Similarly, assume a model that is trained on an external data source. This means that it will recognise the the cases where some 'h' were replaced with 'b', but not recognise that 'Reed' has become 'Reeal'. Hence, the external consistency score would be 58/62 = 0.935.

# Application

**(Keli adds application in OCR tasks)**

[Internal consistency: ground truth of the text and the text that are aligned with ground truth]

# Discussion

**(Keli organizes these text to illustrate limitations and future works)**

Internal validity vs external- one is their own words the other is a general set of words.

In the same way that precision and recall provide important measures…

However, n-gram models do not take the linguistic structure of language into account. For instance, @rosenfeld2000two [p. 1] discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'. The prediction of the next word is based on only a few preceding words and broader context is not taken into account. Hence next-word prediction using n-gram based language models can be limited.

Here think of a two-gram involving the word 'good' 'good morning'. At scale these can identify missing or unusual words, and work quickly, but they lack nuance. For instance, an equally reasonable two-gram involving the word 'good' is 'good work'. For that reason, we consider pre-trained word embedding models. These include Word2Vec and GloVe, which place each word in a multi-dimensional space such that distance between words can illustrate their relationship. We also consider pre-trained generative models such as GPT-2 and GPT-3 and BERT. GPT-2, GPT-3 and BERT are pre-trained generative unsupervised language models. GPT-2 and GPT-3 are the two generations of the same model that differ in the number of parameters. GPT-2 has 1.5 billion parameters and GPT-3 has more than 175 billion parameters. While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot and is accessed via an API. BERT has 340 millions parameters which is smaller than both GPT-2 and GPT-3. BERT also differs from GPT models because its encoding process is bidirectional instead of unidirectional.

>>>>

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    toc: false
    fig_caption: true
    latex_engine: pdflatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
biblio-style: apalike
title: "aRianna: R package for consistency scores in text data"
thanks: "Thank you to **X, Y and Z** for helpful comments. A Shiny application for interactive and small-scale examples is available: https://kelichiu.shinyapps.io/Arianna/. An R package for larger scale applications is available: https://github.com/RohanAlexander/arianna. Our code and datasets are available: https://github.com/RohanAlexander/consistency_scores_in_text_datasets. Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Ke-Li Chiu
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: University of Toronto
abstract: "In this paper, we introduce a process to clean the text extracted from PDFs using n-gram models that is widely used in statistical natural language processing. Our approach compares originally extracted text with the text generated from, or expected by, these methods using earlier text as stimulus. To guide this process, we introduce the notion of a consistency score, which refers to the proportion of text that is unchanged by the model. This is used to monitor changes during the cleaning process, and to compare the messiness of different texts. We illustrate our process on text from the book Jane Eyre and introduce both a Shiny application and an R package to make our process easier for others to adopt."
keywords: "text-as-data; natural language processing; quantitative analysis."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
# spacing: double
endnote: no
graphics: yes
---

@@

# 1. Introduction

@@

We employ n-gram models to calculate consistency scores and generate word candidates for text corrections. The n-gram approach involves identifying two, three, or more, words that are commonly found together. In our application, we use the sequence of three words, or tri-grams, to calculate consistency scores and generate word candidates for text corrections. By using tri-grams, we identify the first two words in the sequence and evaluate if the last word in the sequence is "as expected" by comparing it to the corresponding tri-gram in a comparison dataset. The comparison dataset can be internal, meaning that the dataset is the source where the text being evaluated is extracted from. On the other hand, a comparison dataset is external when the dataset is from a different source than where the text being evaluated is from, and is a larger and more universal dataset that holds generalizability. The external dataset we use in the application is constructed using a subset of text data sourced from Open WebText Corpus by @Gokaslan2019OpenWeb. Since the content of Open WebText Corpus is from the web, the variety of the text data is promising for meeting the generalizability and universality an external dataset should possess.

In this paper, we introduce the `aRianna` R Package with the following structure: In Section 2, we provide an overview of language model and n-grams. In section we deconstruct the package functions and provide demonstration example of the package usage. Finally, we discuss the limitation and the future directions of aRianna in Section 4. Beyond this paper, we construct a Shiny app available at: https://kelichiu.shinyapps.io/aRianna/. That app computes internal and external consistency scores for corpus excerpts, and have developed an R Package, `aRianna`, that allows our approach to be used on larger datasets, which is available at: https://github.com/RohanAlexander/aRianna.

# 2. Background

@@

# 3. The aRianna R Package

## Package Dependencies

There are a variety of ways to implement an n-gram model within R @citeR including using R packages such as `Quanteda` [@citequanteda], `tidyText` [@citetidytext] and `tm` [@citetm]. In our package, we used the `Quanteda` R package because it has a comprehensive set of functions for conducting text analysis. We also employed `dplyr` [@2018dplyr] for data frame manipulation and `tibble` [@citeTibble] for transforming data frames to the tibble format:

@@

## Function to Make the Internal DataSet

To obtain the internal consistency score, we will construct a internal consistency dataset from a body of text that is the source of the text being evaluated. The function `make_internal_consistency_dataset` is created for this purpose. The function turns the body_of_text into an internal consistency dataset which the text being evaluated is compared with. To do so, the function first transforms the body of text into word tokens with the punctuation removed and the letters turned to lowercase:

@@

Finally, the `make_internal_consistency_dataset` function splits the tri-grams to two parts: the previous two words and the last word. In the end, a tibble is created with each tri-gram containing its original three-word sequence, the first two words, and the last word:

@@

## Function to Generate Internal Consistency Score

After we have the internal consistency dataset, we need a function to compare the text being evaluated and the internal dataset in order to retrieve the internal consistency score — the `generate_internal_consistency_score` function. Similar to the `make_internal_consistency_dataset` function, the function first transforms the text being evaluated into tri-gram tokens, then splits the tri-grams and preserves them into a tibble:

@@

Next, the function combines the text tokens tibble and the internal consistency tibble by the `left_join()` function of `dplyr`. The `left_join()` function returns all the rows from the text tokens tibble and all the columns from both tibbles. By combining the two tibbles, the function generates two additional columns — the `last_word` column are the last words of all tri-grams, and the `last_word_expected` contains the words that are in the internal consistency dataset. By comparing the words in `last_word` and `last_word_expected`, we can identify the "unexpected" words. If a last word is in `last_word` but not `last_word_expected`, it is regarded as an unexpected word and the corresponding word in `last_word_expected` serves as a replacement candidate:

@@

## Function to Generate External Consistency Score

The `generate_external_consistency_score` function works identically with the `generate_internal_consistency_score`. The only difference is that it compares the text being evaluated with a external consistency dataset that is larger and more general. The external consistency dataset we employed in the application is constructed using a subset text data from Open WebText Corpus by @Gokaslan2019OpenWeb.

## Demonstration

To install the package and load the library:

```{r, eval=FALSE}
devtools::install_github("RohanAlexander/arianna")
library(aRianna)
```

@@

```{r, eval=FALSE}
aRianna::generate_internal_consistency_score(
  text_to_evaluate, internal_consistency_dataset)

@@

## $`unexpected words`
## # A tibble: 11 x 4
##    first_words last_word last_word_expected as_expected
##    <chr>       <chr>     <chr>              <lgl>      
##  1 there_was   na        a                  FALSE      
##  2 there_was   na        an                 FALSE      
##  3 there_was   na        no                 FALSE      
##  4 there_was   na        not                FALSE      
##  5 there_was   na        nothing            FALSE      
##  6 there_was   na        of                 FALSE      
##  7 there_was   na        one                FALSE      
##  8 there_was   na        only               FALSE      
##  9 there_was   na        plenty             FALSE      
## 10 there_was   na        someone            FALSE      
## 11 there_was   na        something          FALSE

@@

# 4. Discussion

Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class]. As demonstrated in our application, we use a tri-gram based model to calculate internal and external consistency scores and generate a list of words as text correction candidates. However, n-gram models do not take the linguistic structure of language into account. For instance, @rosenfeld2000two [p. 1] discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'. The prediction of the next word is based on only a few preceding words and broader context is not taken into account. Hence next-word prediction using n-gram based language models can be limited. Here think of a two-gram involving the word 'good' 'good morning'. At scale these can identify missing or unusual words, and work quickly, but they lack nuance. For instance, an equally reasonable two-gram involving the word 'good' is 'good work'. However, because n-gram models does not take the surrounding text into account, it has not capacity of judging if 'morining' or 'work' is to be suggested as the next word after 'good'.

In future works, we intend to investigate in more advanced language models such as word embeddings and Transformers. We consider pre-trained word embedding models including Word2Vec [@mikolov2013efficient] [@mikolov2013distributed] and GloVe [@pennington2014glove], which place each word in a multi-dimensional space such that distance between words can illustrate their relationship [@bengio2003neural]. This approach of representing words in vectors has a long history, but the implementation of @bengio2003neural has become the foundation for much subsequent work. Here, terms are encoded into feature vectors that represent different aspects of the term, and so each term is represented by a position in the vector space.

We also consider incorporating generative pre-trained transformer models such as GPT-2 [@radford2019language] and GPT-3 [@brown2020language] and BERT [@devlin2018bert]. GPT-2, GPT-3 and BERT are pre-trained language models that are built based on the Transformer architecture [@vaswani2017attention]. Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based and CNN-based models in computational efficiency [@vaswani2017attention]. Since the emergence of the Transformer model, most of the representative pre-trained models are built based on this architecture [@hanretty2018comparing]. Therefore, we can say that Transformer marks a new era of language models. In our future works, it will be important for us to implement different language models compare their effectiveness in the tasks of generating consistency score and providing suggested text corrections and improve the `aRianna` package.

\newpage
